\section{Related Work}

%Burstiness on topic model:
%Modeling Word Burstiness Using the Dirichlet Distribution (DCM)
%Accounting for Burstiness in Topic Models (DCMLDA)
%Proposal of a-MMSB in : Scalable Inference of Overlapping Communities with high diagonal only...

%to read: Stochastic blockmodels and community structure in networks

Recently, the MMSB \cite{MMSB} and ILFM \cite{ILFM} models have been successfully used for link predictions and structure discovery in social networks. Several extensions, has been proposed. In \cite{AMMSB}, they proposed a adaptation of a MMSB model that scale well using stochastic variational inference and use it for the discovery of overlapping communities in big networks (millions of nodes). They constrained the weight matrix to have weights in the diagonal and a fixed small value elsewhere and named it a-MMSB where "a" stand for assortative. The MMSB model has also been extended to a nonparametric dynamic version to handle temporal networks \cite{fan2015dynamic}. ILFM has also been extended in several way; to handle non-negative weights in \cite{IMRM} and within a more subtle latent feature structure in \cite{ILAM}. Nevertheless the characterization of models with regards to the properties for networks remains to be explored \cite{jacobs2014unified}. Theoretical results on the structure of random graph under exchangeability assumptions are however well reported in \cite{orbanz2015bayesian}, although the characterization of social networks properties within a Bayesian framework is not covered. In \cite{hoff2008modeling}, the role of exchangeability was first pointed out and its relation to the data representation where they provide a comparative study of the homophily and structural equivalence in different model settings, but without a formal definition of both. An other important mention of the exchangeability assumptions, was done in \cite{bickel2009nonparametric} where they characterize the asymptotic comportment of the likelihood in the blockmodel settings, showing that the inference of the likelihood lead to a consistent estimation of the community structure. But they only cover the case a single membership.

